Propositions as Sessions, Mechanically

### 2015-04-10 paper/formalisation.tex
Added select rules regarding GV and some initial discussion.

%\input{gv}

The typing judgements for the functional language GV are shown in
Figure~\ref{??}. Note that I have altered the $n$-ary branch and choice
constructs to the restricted binary versions in order to avoid tricky
formulations. The change is not restrictive however, since CP itself provides
only binary versions of plus ($\oplus$) and with ($\with$)
constructs. Explicit limiting and unlimiting of abstraction was chosen to
enable the inclusion of these rules in the translation from GV to CP, which is
performed by induction on GV terms.

@@

\subsection{CPS translation GV to CP}

require type inference on gv terms to construct corresponding cp terms but we
are assured that this is okay since the theorem only provides a well-typed gv
term in a context that is well-defined

type inference of gv terms required cluttered definition of translation with
impossible error cases. avoided a dependently typed solution due to
difficulties encountered earlier in the development. james mckinna et
al. suggest encoding the successful cases in an inductive type and then
defining the function to act on these cases. will this really work? this means
a graph of positive cases for types then a function definition for it too this
seems rather pathetic because i have to define th einference on terms and it
seems wrong i have then to handle environments but also nonenvironments i
can't remember what they are called unbound variables that is it they have to
be handled as well so we have all this clutter and the types too that is the
important part but the environments then have to be handled linearly
unfortunately very bad because we could just use the well typed term relation
for most of this headachey breaky stuff i'm not sure about that but what is
can i say so for example lets see see no i also have to thing about the
general guidelines i can be giving for this section so we have something to
say about guidelines do we yes of course we have can say that out of this we
realised that it is important for a representation to be good enough to
succintly express that which we wish to say in particular the graph relation
the inductive type that we define propositionally is better one would think
than the other way the monadic possible failure way is not producing a valid
thing that we can inductively induct on so we concern ourself with the proof
of translation preserving the typed term and i think it is important to
remember the problem we had with the inductive hypothesis we had something do
with translating the inner term of a weakening term in a larger context than
the one we had available and that was to do with the destruction of the term
inference thing i think no no it was something else we had the inner term and
we had it well typed in an environment that was smaller sure and then our
thing our function was saying we had some environment that typed both the
inner term and the outer weakening term but we know that is wrong so wrong so
here is perhaps where the inductive type graph relation thing wins out in that
we can have a hypothesis for the inner term which is sufficiently small for
the environment and we pattern match not only on the term but also on the
environment yes that is we will do because it seems to work only when we get
to the messy two the messy two constructor caases will this become a bigg
problem inthat we will then have all this stuff about environment being split
and needing to be recombined in the composite process that is a term not a
process but the point still stands it will need to be about the term and the
splitting is really a duplication of the unfortunately the well typed term
relation but otherwise i cannot think how to reuse or obtain that information
for the translation to cp

Translation uses the coq-ext-lib from michela for monadic notation

Currently the formalisation does not handle this. Future work.

\todo{how might we make progress on this?}

@@

\textbf{Modularised development} Not only is modular development an important
programming practice in general, but in the context of theorem proving it has
the advantage of reducing interactive load time. Executing proof scripts can
be time-consuming especially if those scripts make frequent use of decision
procedures performing proof search, then the execution time is proportional to
the size of the proof context and complexity of the goal. There were cases in
which the system crashed unexpectedly during interactive mode and it is
conceivable that having many large proofs in-memory could be a contributing
factor. For these reasons, the development is partitioned into GV and CP
sections each with sub-divided into definition, infrastructure and typing
modules (CP also has an additional module for reduction lemmas). The CPS
translation is also defined separately. Load time is considerably improved and
the consequences of a crash are mitigated.

\textbf{Representation} It is important not to go overboard in the
when developing not to use dependent types all the time without working out if
it will be to the advantage of the prover use dependent types and other power
with care for example we could don't want to use mutual inductive definitions
for the types construct since this is overly complicated for what we wish to
achieve and we don't want to index types with kinds since this complicates
further definitions it is about trying to use dependent types responsibly
without using them in an uncritical or thinking manner with great power comes
great responsibility

\textbf{Library reuse} One of the design goals of the development was to reuse
existing libraries wherever possible. Two such instances proved extremely
useful, first is the Metatheory library as already mentioned. Second, is the
library of tactics provided by the Software Foundations online book~\cite{SF}.
The tactics helped in many ways including simplifying introduction of
assertions, providing convenient shorthands and ... Traditionally, assertions
in Coq have to be explicitly written within a proof leading to the use of
empheremal names introduced by prover. If the proof of the assertion follows
from an already proved lemma and some assumptions in the context this is
unnecessary verbage and this is what the library allows to omit \todo{example
  of using the forwards tactic} an example of the so called forwards tactic is
given below the $\sim$ is a a shorthand for immediately applying \coqe$auto$
to all subgoals generated by the tactic.

\textbf{Linear environments} Linearity create issues with typing some
examples; the splitting of the environment was non-deterministic so required
side-conditions to enforce a certain splitting of binders. For example the
rule for associativity of cut. \todo{how to handle? is there a better way?}

>>>>

\input{gv}

The typing judgements for the functional language GV are shown in
Figure~\ref{fig:gv}. Note that I have altered the $n$-ary branch and choice
constructs to the restricted binary versions in order to avoid tricky
formulations. The change is not restrictive however, since CP itself provides
only binary versions of plus ($\oplus$) and with ($\with$)
constructs. Explicit limiting and unlimiting of abstraction was chosen to
enable the inclusion of these rules in the translation from GV to CP, which is
performed by induction on GV terms. \todo{could take about opening the
  multiple binder let. multi but not as in literature}

@@

\subsection{CPS translation from GV to CP}

Following the original work by \citeauthor{Wadler:2014}, the translation from
GV to CP requires defining translations on types and terms of GV, which for
terms is in continuation-passing style.

\subsubsection{Types}

The translation of from GV types to CP propositons is shown in Figure~\ref{??}
\todo{Coq version or regular version?}. Translation of session types is
noteworthy for taking a GV session type to its notional dual in CP, reflecting
the semantic difference between channels in GV and those in CP.

\subsubsection{Terms}

Due to the locally nameless representation, the translation of GV terms
required type inference of terms to construct corresponding CP terms in order
to specify the propositions. So, the translation takes the environment, in
which the term was typed, as an argument to query the type of free
variables. Unfortunately, this query returns a type of \coqe!option typ!
(\coqe!option! is equivalent to \tm{Maybe} in Haskell) since it may not be
found, but in fact since only well-typed GV terms are translated it is
guaranteed to return something! Therefore, the definition is unnecessarily
cluttered with impossible error cases.

The term translation was initially designed using a monadic approach to
account for the possible error cases in retreiving names from the
environment. The coq-ext-lib
library~\footnote{https://github.com/coq-ext-lib/coq-ext-lib} provides
definitions for monad, functor and applicative type classes and associated
instances for basic Coq types. It is similar in style to Haskell's support for
these objects. Unfortunately, this definition did not yield a suitable
function to prove the translation well-typed because the function needs to be
able to operate structurally on not only terms, but also the environment and
the processes being constructed.

An alternative formulation, which avoids the need to handle error cases, is to
first define only the positive cases using an inductive type, the so-called
\textit{graph relation}, and then define a dependent function which only
considers the positive cases within the relation~\cite{McKinna:2009}. However,
several issues were encountered when attempting to specify the
translation. The graph relation requires much of the clutter present in both
the \coqe!cp_rule! and \coqe!wt_tm! definition such as linear enviornments and
cofinite quanitifcation.

i'm not sure about that but what is can i say so for example lets see see no i
also have to thing about the general guidelines i can be giving for this
section so we have something to say about guidelines do we yes of course we
have can say that out of this we realised that it is important for a
representation to be good enough to succintly express that which we wish to
say in particular the graph relation the inductive type that we define
propositionally is better one would think than the other way the monadic
possible failure way is not producing a valid thing that we can inductively
induct on so we concern ourself with the proof of translation preserving the
typed term and i think it is important to remember the problem we had with the
inductive hypothesis we had something do with translating the inner term of a
weakening term in a larger context than the one we had available and that was
to do with the destruction of the term inference thing i think no no it was
something else we had the inner term and we had it well typed in an
environment that was smaller sure and then our thing our function was saying
we had some environment that typed both the inner term and the outer weakening
term but we know that is wrong so wrong so here is perhaps where the inductive
type graph relation thing wins out in that we can have a hypothesis for the
inner term which is sufficiently small for the environment and we pattern
match not only on the term but also on the environment yes that is we will do
because it seems to work only when we get to the messy two the messy two
constructor caases will this become a bigg problem inthat we will then have
all this stuff about environment being split and needing to be recombined in
the composite process that is a term not a process but the point still stands
it will need to be about the term and the splitting is really a duplication of
the unfortunately the well typed term relation but otherwise i cannot think
how to reuse or obtain that information for the translation to cp

Roughly, the theorm aiming to prove is:

@@

\textbf{Modularised development.} Not only is modular development an important
programming practice in general, but in the context of theorem proving it has
the advantage of reducing interactive load time. Executing proof scripts can
be time-consuming especially if those scripts make frequent use of decision
procedures performing proof search, then the execution time is proportional to
the size of the proof context and complexity of the goal. There were cases in
which the system crashed unexpectedly during interactive mode and it is
conceivable that having many large proofs in-memory could be a contributing
factor. For these reasons, the development is partitioned into GV and CP
sections each with sub-divided into definition, infrastructure and typing
modules (CP also has an additional module for reduction lemmas). The CPS
translation is also defined separately. Load time is considerably improved and
the consequences of a crash are mitigated.

\textbf{Representation.} It is important not to go overboard in the
when developing not to use dependent types all the time without working out if
it will be to the advantage of the prover use dependent types and other power
with care for example we could don't want to use mutual inductive definitions
for the types construct since this is overly complicated for what we wish to
achieve and we don't want to index types with kinds since this complicates
further definitions it is about trying to use dependent types responsibly
without using them in an uncritical or thinking manner with great power comes
great responsibility

\textbf{Library reuse.} One of the design goals of the development was to
reuse existing libraries wherever possible. Two such instances proved
extremely useful, first is the Metatheory library as already
mentioned. Second, is the library of tactics provided by the Software
Foundations online book~\cite{SF}.  The tactics helped in many ways including
simplifying introduction of assertions, providing convenient shorthands and
... Traditionally, assertions in Coq have to be explicitly written within a
proof leading to the use of empheremal names introduced by prover. If the
proof of the assertion follows from an already proved lemma and some
assumptions in the context this is unnecessary verbage and this is what the
library allows to omit \todo{example of using the forwards tactic} an example
of the so called forwards tactic is given below the $\sim$ is a a shorthand
for immediately applying \coqe$auto$ to all subgoals generated by the tactic.

\textbf{Linear environments.} Linearity create issues with typing some
examples; the splitting of the environment was non-deterministic so required
side-conditions to enforce a certain splitting of binders. For example the
rule for associativity of cut. \todo{how to handle? is there a better way?}
